Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-06-22 lqg-strings.tex
made some changes to section IV

% Mar 25, 2023
% Added parskip package for spacing between paragraphs.
\usepackage{parskip}
% \setlength{\parindent}{15pt}

@@

Equation (\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal factor $\e ^{\phi (x)}$ appearing in $F^{-}$. By using the ansatz in eq (\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal factor to fix the second term as $- m^2 X^{\mu}$, with $- m^2 = p^2 - E^2$ on-shell and this, in turn, implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

\begin{equation} \label{eqn:F-minus-exp}
	F^{-} = A \e ^{i (E \tau + p \sigma)} , \ A \in \C
\end{equation}
\begin{equation} \label{eqn:phi}
	\e ^{\phi (x)} = \sqrt{\frac{k \Delta}{2 \left( 1 - A \e ^{i (E \tau + p \sigma)} \right)}} .
\end{equation}
At first glance, it seems problematic that we have a complex conformal factor, but the complex structure comes from our ansatz for the embedding fields, and when we choose either the real or imaginary part of the latter it should also pick the respective component of $\e ^{\phi}$. With this choice, we thus turned eq (\ref{eqn:X-eom-simplified}) into a Klein-Gordon equation
\begin{equation} \label{eqn:X-Klein-Gordon-eqn}
	\left( \partial _{\sigma} ^2 - \partial _{\tau} ^2 \right) X^{\mu} - \mu ^2 X^{\mu} = 0 .
\end{equation}
The imposition that
\begin{equation} \label{eqn:mu-constraint-1}
	i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) = - \mu ^2
\end{equation}
can be used to find the explicit dependence of $\mu ^2$ on the quantum geometry parameters $k \Delta$ as in
\begin{multline} \label{eqn:mu-constraint-final}
	\mu ^2 = \frac{i}{\dfrac{2 \e ^{2 \phi}}{k \Delta} - 1} \big( p \del _{\sigma} \ln \left( \e^{2 \phi} / \Delta \right) - \\
	- E \del _{\tau} \ln \left( \e^{2 \phi} / \Delta \right) \big) .
\end{multline}

@@

From our analysis of the inverse-area corrected Polyakov action, we got $D$ finite-space one-dimensional Klein-Gordon equations (\ref{eqn:X-Klein-Gordon-eqn}), with constraints given by eq\eqref{eqn:inverse-area-em-tensor}. The constraints are nothing new, they're precisely the same as the ones of regular String Theory. What's new in our analysis is the $\mu ^2$ term, which we will assume to be constant for a first analysis, making it so the solution is given by Fourier transform
\begin{multline} \label{eqn:KG-Polyakov-solution-initial}
	X^{\mu} (\tau , \sigma) = \sqrt{\frac{\alpha '}{2}} \sum _{p \in \mathbb{Z}} \frac{1}{2 E_{p}} \bigg( \alpha ^{\mu} _{p} \e ^{i (E_{p} \tau - p \sigma)} + \\
	+ \beta ^{\mu} _{p} \e ^{- i (E_{p} \tau - p \sigma)} \bigg) ,
\end{multline}
with $E_{p} = \sqrt{p^2 + \mu ^2}$. From here on, following an analogous procedure to the one found in \cite{Tong2009Lectures}, reality of $X^{\mu}$ implies $(X^{\mu})^{*} = X^{\mu}$, from which we get
\begin{equation} \label{eqn:reality-consequence}
	\beta ^{\mu} _{n} = (\alpha ^{\mu} _{n})^{*} .
\end{equation}

@@

Looking closer at our Virasoro-like \textquotedblleft modes", in particular the $n = p$ condition, we get the mass of the string in terms of its vibrational modes and center of mass position
\begin{multline} \label{eqn:string-rest-mass-classical-1}
	M^2 = - p^{\mu} p_{\mu} = \frac{4}{\alpha '} \Bigg( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 \widetilde{L}_{n , n} \right) - \\
	- \frac{2 \mu ^2}{\alpha '} x^{\mu} x_{\mu} \Bigg) =
\end{multline}
\begin{multline} \label{eqn:string-rest-mass-classical-2}
	= \frac{4}{\alpha '} \Bigg( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 \alpha ^{\mu} _{n} (\alpha ^{\nu} _{n})^{*} \eta _{\mu \nu} \right) - \\
	- \frac{2 \mu ^2}{\alpha '} x^{\mu} x_{\mu} \Bigg) .
\end{multline}

@@

We start by calculating the canonical momentum density conjugate to the fields $X^{\mu} (x)$, $\Pi ^{\nu} (x)$, given by
\begin{multline} \label{eqn:canonical-momentum-density}
	\Pi ^{\nu} (x) = F^{-} \frac{i}{8 \pi \alpha '} \sum _{n} \bigg( \alpha ^{\nu} _{n} \e ^{i (E_{n} \tau - n \sigma)} - \\
	- (\alpha ^{\nu} _{n})^{*} \e ^{- i (E_{n} \tau - n \sigma)} \bigg)
\end{multline}
where
\begin{multline} \label{eqn:F-minus-full-sum}
	F^{-} = 1 - \frac{k \Delta}{2 \e ^{2 \phi (x)}} = \\
	= \sum _{l} \left( A_{l} \e ^{i (E_{l} \tau + l \sigma)} + (A_{l})^{*} \e ^{- i (E_{l} \tau + l \sigma)} \right) ,
\end{multline}
and then proceed to use the equal-time Poisson bracket relation at $\tau = 0$
\begin{equation} \label{eqn:canonical-bracket-relation}
	\left\{ X^{\mu} (0 , \sigma) , \Pi ^{\nu} (0 , \sigma ') \right\} = \delta (\sigma - \sigma ') \eta ^{\mu \nu}
\end{equation}
to calculate the bracket relations between the expansion modes $\alpha ^{\mu} _{n}$ and $(\alpha ^{\nu} _{p})^{*}$. After using the freedom we still have in choosing the expansion modes of $F^{-}$ to gauge fix it to satisfy $\mathfrak{Re} (A_{0}) = - 2 \alpha '$ and realizing that $\alpha ^{\mu} _{- n} = \alpha ^{\mu} _{n}$, we get the desired
\begin{subequations}
	\begin{align}
		\left\{ \alpha ^{\mu} _{n} , \alpha ^{\nu} _{p} \right\}       & = 0 = \left\{ (\alpha ^{\mu} _{n})^{*} , (\alpha ^{\nu} _{p})^{*} \right\} \label{eqn:expansion-modes-bracket-0} \\
		\left\{ \alpha ^{\mu} _{n} , (\alpha ^{\nu} _{p})^{*} \right\} & = - i E_{n} \delta _{|n| , |p|} \eta ^{\mu \nu} , \label{eqn:expansion-modes-bracket-1}
	\end{align}
\end{subequations}
from which we can normalize the expansion modes to get harmonic oscillator modes and bracket relations
\begin{subequations}
	\begin{align}
		a^{\mu} _{n}       & := \frac{1}{\sqrt{E_{n}}} \alpha ^{\mu} _{n} \label{eqn:normalized-oscillator-modes}         \\
		(a^{\mu} _{n})^{*} & := \frac{1}{\sqrt{E_{n}}} (\alpha ^{\mu} _{n})^{*} \label{eqn:normalized-oscillator-modes-*}
	\end{align}
\end{subequations}
\begin{equation} \label{eqn:harmonic-bracket-relations}
	\left\{ a^{\mu} _{n} , (a^{\nu} _{p})^{*} \right\} = - i \delta _{|m| , |l|} \eta ^{\mu \nu} .
\end{equation}
The bracket relations for $\alpha ^{\mu} _{0}$ can be used to get the expected bracket relations between $x^{\mu}$ and $p^{\nu}$
\begin{align}
	\left\{ x^{\mu} , x^{\nu} \right\} & = 0 = \left\{ p^{\mu} , p^{\nu} \right\} \label{eqn:x-x-p-p-brackets} \\
	\left\{ x^{\mu} , p^{\nu} \right\} & = \eta ^{\mu \nu} . \label{eqn:x-p-bracket-relations}
\end{align}

@@

Having the bracket relations between the vibrational modes of the string, we can now calculate the constraint algebra, which we shall refer to as the Virasoro-Klein-Gordon Algebra. For simplicity, since $E_{n} = \sqrt{n^2 + \mu ^2} > 0$, we shall work with the normalized harmonic oscillator modes $a^{\mu} _{n}$ instead of the expansion modes $\alpha ^{\mu} _{n}$. The algebra is given by
\begin{subequations}
	\begin{equation} \label{eqn:L-brackets-0}
		\left\{ L_{n , m} , L_{k , p} \right\} = 0
	\end{equation}
	\begin{equation} \label{eqn:L-brackets-1}
		\left\{ L_{n , m} , \widetilde{L}_{k , p} \right\} = - i \left( L_{m , k} \delta _{|n| , |p|} + L_{n , k} \delta _{|m| , |p|} \right)
	\end{equation}
	\begin{equation} \label{eqn:L-brackets-2}
		\left\{ \widetilde{L}_{n , m} , \widetilde{L}_{k , p} \right\} = - i \left( \widetilde{L}_{k , m} \delta _{|n| , |p|} - \widetilde{L}_{n , p} \delta _{|m| , |k|} \right) .
	\end{equation}
\end{subequations}

>>>>

% Mar 25, 2023
% Added parskip package for spacing between paragraphs.
\usepackage{parskip}
\setlength{\parindent}{15pt}

@@

Equation (\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal factor $\e ^{\phi (x)}$ appearing in $F^{-}$. By using the ansatz in eq (\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal factor to fix the second term as $- \mu ^2 X^{\mu}$, with $- \mu ^2 = p^2 - E^2$ on-shell and this, in turn, implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

\begin{equation} \label{eqn:F-minus-exp}
	F^{-} = A \e ^{i (E \tau + p \sigma)} , \ A \in \C
\end{equation}
\begin{equation} \label{eqn:phi}
	\e ^{\phi (x)} = \sqrt{\frac{k \Delta}{2 \left( 1 - A \e ^{i (E \tau + p \sigma)} \right)}} .
\end{equation}
It may seem problematic to have a complex conformal factor, but the complex structure comes from our ansatz for the embedding fields, and when we choose either the real or imaginary part of the latter it should also pick the respective component of $\e ^{\phi}$. With this choice, we thus turned eq (\ref{eqn:X-eom-simplified}) into a Klein-Gordon equation
\begin{equation} \label{eqn:X-Klein-Gordon-eqn}
	\left( \partial _{\sigma} ^2 - \partial _{\tau} ^2 \right) X^{\mu} - \mu ^2 X^{\mu} = 0 .
\end{equation}
The imposition that
\begin{equation} \label{eqn:mu-constraint-1}
	i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) = - \mu ^2
\end{equation}
can be used to find the explicit dependence of $\mu ^2$ on the quantum geometry parameters $k \Delta$ as
\begin{multline} \label{eqn:mu-constraint-final}
	\mu ^2 = \frac{i}{\dfrac{2 \e ^{2 \phi}}{k \Delta} - 1} \big( p \del _{\sigma} \ln \left( \e^{2 \phi} / \Delta \right) - \\
	- E \del _{\tau} \ln \left( \e^{2 \phi} / \Delta \right) \big) .
\end{multline}

@@

From our analysis of the inverse-area corrected Polyakov action, we got $D$ finite-space one-dimensional Klein-Gordon equations (\ref{eqn:X-Klein-Gordon-eqn}), with constraints given by eq \eqref{eqn:inverse-area-em-tensor}. The constraints are nothing new, they're the same as the ones of regular String Theory. What's new in our analysis is the $\mu ^2$ term, which we take as being constant for a first analysis, making it so the solution is given by Fourier transform
\begin{multline} \label{eqn:KG-Polyakov-solution-initial}
	X^{\mu} (\tau , \sigma) = \sqrt{\frac{\alpha '}{2}} \sum _{p \in \mathbb{Z}} \frac{1}{2 E_{p}} \bigg( \alpha ^{\mu} _{p} \e ^{i (E_{p} \tau - p \sigma)} + \\
	+ \beta ^{\mu} _{p} \e ^{- i (E_{p} \tau - p \sigma)} \bigg) ,
\end{multline}
with $E_{p} = \sqrt{p^2 + \mu ^2}$. From here on, following an analogous procedure to the one found in \cite{Tong2009Lectures}, reality of $X^{\mu}$ implies $(X^{\mu})^{*} = X^{\mu}$, from which we get
\begin{equation} \label{eqn:reality-consequence}
	\beta ^{\mu} _{n} = (\alpha ^{\mu} _{n})^{*} .
\end{equation}

@@

Looking closer at our constraints \textquotedblleft modes", in particular the $n = p$ condition, we get the mass of the string in terms of its vibrational modes and center of mass position
\begin{multline} \label{eqn:string-rest-mass-classical-1}
	M^2 = - p^{\mu} p_{\mu} = \frac{4}{\alpha '} \Bigg( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 \widetilde{L}_{n , n} \right) - \\
	- \frac{2 \mu ^2}{\alpha '} x^{\mu} x_{\mu} \Bigg) =
\end{multline}
\begin{multline} \label{eqn:string-rest-mass-classical-2}
	= \frac{4}{\alpha '} \Bigg( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 \alpha ^{\mu} _{n} (\alpha ^{\nu} _{n})^{*} \eta _{\mu \nu} \right) - \\
	- \frac{2 \mu ^2}{\alpha '} x^{\mu} x_{\mu} \Bigg) .
\end{multline}

@@

We start by calculating the canonical momentum density conjugate to the fields $X^{\mu} (x)$, $\Pi ^{\nu} (x)$, given by
\begin{multline} \label{eqn:canonical-momentum-density}
	\Pi ^{\nu} (x) = F^{-} \frac{i}{8 \pi \alpha '} \sum _{n} \bigg( \alpha ^{\nu} _{n} \e ^{i (E_{n} \tau - n \sigma)} - \\
	- (\alpha ^{\nu} _{n})^{*} \e ^{- i (E_{n} \tau - n \sigma)} \bigg)
\end{multline}
where
\begin{multline} \label{eqn:F-minus-full-sum}
	F^{-} = 1 - \frac{k \Delta}{2 \e ^{2 \phi (x)}} = \\
	= \sum _{l} \left( A_{l} \e ^{i (E_{l} \tau + l \sigma)} + (A_{l})^{*} \e ^{- i (E_{l} \tau + l \sigma)} \right) ,
\end{multline}
and then proceed to use the equal-time Poisson bracket relation at $\tau = 0$
\begin{equation} \label{eqn:canonical-bracket-relation}
	\left\{ X^{\mu} (0 , \sigma) , \Pi ^{\nu} (0 , \sigma ') \right\} = \delta (\sigma - \sigma ') \eta ^{\mu \nu}
\end{equation}
to calculate the bracket relations between the expansion modes $\alpha ^{\mu} _{n}$ and $(\alpha ^{\nu} _{p})^{*}$. After using the freedom we still have in choosing the expansion modes of $F^{-}$ to gauge fix it to satisfy $\mathfrak{Re} (A_{0}) = - 2 \alpha '$ and realizing that $\alpha ^{\mu} _{- n} = \alpha ^{\mu} _{n}$, we get the expected
\begin{subequations}
	\begin{align}
		\left\{ \alpha ^{\mu} _{n} , \alpha ^{\nu} _{p} \right\}       & = 0 = \left\{ (\alpha ^{\mu} _{n})^{*} , (\alpha ^{\nu} _{p})^{*} \right\} \label{eqn:expansion-modes-bracket-0} \\
		\left\{ \alpha ^{\mu} _{n} , (\alpha ^{\nu} _{p})^{*} \right\} & = - i E_{n} \delta _{|n| , |p|} \eta ^{\mu \nu} , \label{eqn:expansion-modes-bracket-1}
	\end{align}
\end{subequations}
from which we can normalize the expansion modes to get harmonic oscillator modes and bracket relations
\begin{subequations}
	\begin{align}
		a^{\mu} _{n}       & := \frac{1}{\sqrt{E_{n}}} \alpha ^{\mu} _{n} \label{eqn:normalized-oscillator-modes}         \\
		(a^{\mu} _{n})^{*} & := \frac{1}{\sqrt{E_{n}}} (\alpha ^{\mu} _{n})^{*} \label{eqn:normalized-oscillator-modes-*}
	\end{align}
\end{subequations}
\begin{equation} \label{eqn:harmonic-bracket-relations}
	\left\{ a^{\mu} _{n} , (a^{\nu} _{p})^{*} \right\} = - i \delta _{|n| , |p|} \eta ^{\mu \nu} .
\end{equation}
The bracket relations for $\alpha ^{\mu} _{0}$ can be used to get the expected bracket relations between $x^{\mu}$ and $p^{\nu}$
\begin{align}
	\left\{ x^{\mu} , x^{\nu} \right\} & = 0 = \left\{ p^{\mu} , p^{\nu} \right\} \label{eqn:x-x-p-p-brackets} \\
	\left\{ x^{\mu} , p^{\nu} \right\} & = \eta ^{\mu \nu} . \label{eqn:x-p-bracket-relations}
\end{align}

@@

Having the bracket relations between the vibrational modes of the string, we can now calculate the constraints algebra, which we shall refer to as the Virasoro-Klein-Gordon Algebra. For simplicity, since $E_{n} = \sqrt{n^2 + \mu ^2} > 0$, we shall work with the normalized harmonic oscillator modes $a^{\mu} _{n}$ instead of the expansion modes $\alpha ^{\mu} _{n}$. Then, the algebra in terms of said normalized modes is given by
\begin{subequations}
	\begin{equation} \label{eqn:L-brackets-0}
		\left\{ L_{n , m} , L_{k , p} \right\} = 0
	\end{equation}
	\begin{equation} \label{eqn:L-brackets-1}
		\left\{ L_{n , m} , \widetilde{L}_{k , p} \right\} = - i \left( L_{m , k} \delta _{|n| , |p|} + L_{n , k} \delta _{|m| , |p|} \right)
	\end{equation}
	\begin{equation} \label{eqn:L-brackets-2}
		\left\{ \widetilde{L}_{n , m} , \widetilde{L}_{k , p} \right\} = - i \left( \widetilde{L}_{k , m} \delta _{|n| , |p|} - \widetilde{L}_{n , p} \delta _{|m| , |k|} \right) .
	\end{equation}
\end{subequations}
